### 2023-03-13 paper/FPF-WG1.tex
added abstract

\begin{center}
  {\Large \bf Deep-Inelastic Scattering with TeV Neutrinos at the LHC}
\vspace{1.1cm}

@@

\end{center}
Notes to collect the studies
carried out in WG1 of the Forward
Physics Facility study: {\bf  Neutrino Interactions and Hadronic Structure}

@@

\newline
\newline
We start with a perfect detector.
\newline
\newline
For bins in $x,Q^2,E_{\nu}$, we start with the \href{https://github.com/juanrojochacon/FPF-WG1/blob/main/results/diff_xsecs_a1.txt}{differential cross section},$\frac{d^2\sigma(x,Q^2,E_{\nu}}{dxdQ^2})$ [pb GeV$^{-2}$] as output from {\tt YADISM} and convovle with the \href{https://github.com/KlingFelix/FastNeutrinoFluxSimulation/tree/main/Fluxes}{neutrino flux}, $\frac{dN_{\nu}}{dE_{\nu}}$. So for a detector with $n_T$ nuclear target density and $L_T$ length, we calculate the event rate per bin as

@@

The statistical uncertainty for each bin is then defined by a gaussian distribution with fractional uncertainty $1/\sqrt{N{\rm ev}/{\rm bin}}$, i.e. $\delta N_{\rm ev}/{\rm bin} = \sqrt{N_{\rm ev} /{\rm bin}$
\newline
Following this prescription, we can produce plots which characterize the event rate for a particular neutrino species at a particular experiment.

Results for FASER$\nu$2 can be found \href{https://github.com/juanrojochacon/FPF-WG1/tree/main/results}{here}. The .txt files will display which neutrino species was considered, and also compile experimental geometries and target details in the header. In the subfolder \href{https://github.com/juanrojochacon/FPF-WG1/tree/main/results/Summaries}{summaries}, plots can be found which show the event rate in $x,Q^2$ space, integrating over the entire neutrino spectrum. The event rate for muon neutrinos at FASER$\nu$2 is shown in Fig~\ref{fig:fasernu2_muon}.

@@

We first generate a MC dataset, $D_0$, over this space with $N = 10^7$ samples and calculate $x,Q^2,E_{\nu}$ for each event, removing samples with $y > 1$. We then integrate according to Eq. \ref{MCintegration} to produce a distribution of events with bins in $x,Q^2,E_{\nu}$. We then smear each sample according to a gaussian distribution with widths given by Eq.~\ref{fasernu2systematic_errors} to produce a new data set $D_i$ and repeat to produce $M = 10$ event distributions($M>10$ can later be used, but $M = 10$ appears to be stable). For each bin we take the standard deviation to produce the systematic errors per bin (denoted as "N\textunderscore sys\textunderscore errs" in files named "binned\textunderscore sysevents... .txt" \href{https://github.com/juanrojochacon/FPF-WG1/tree/main/results}{here}).

For FASER${\nu}$2, the systematic erros typically dominate over statsitcal errors, at about $10\%$.

\begin{figure}[h]
    \centering
    \includegraphics[width = 1\textwidth]{error_plot_FASERv2_14.pdf}
    \caption{Event rates with errorbars at FASER$\nu$2 for $\nu_{\mu} N \rightarrow \mu^- X$} summed over $Q^2$. The errobars along the x-axis showing the width of the $x$-bins, and are not showing any type of error.
    \label{fig:my_label}
\end{figure}

@@

The expected impact of the FPF data on global PDF fits is assessed by a profiling procedure~\cite{Paukkunen:2014zia, Schmidt:2018hvu, AbdulKhalek:2018rok, HERAFitterdevelopersTeam:2015cre}, based on minimizing the function
%TODO correlations are also included here, as we hopefully get there in the end -- if not, remove terms involving correlations+pseudodata
\begin{equation}
\chi^2 = 
\sum_{i=1}^{N_{\textrm{bins}}} 
\frac{\left(  \sigma_i^{\textrm{pd}}
            + \Gamma_i^{\alpha,\textrm{pd}}
              b_\alpha^{\textrm{pd}}
            - \sigma_i^{\textrm{th}}
            - \Gamma_i^{\beta,\textrm{th}}
              b_\beta^{\textrm{th}}
     \right)^2
     }{\Delta_i^2}
+ \sum_\alpha (b_\alpha^{\textrm{pd}})^2
+ \sum_\beta  (b_\beta^{\textrm{th}})^2 \, .
\label{eq:profilingchi2}
\end{equation}
Here the pseudodata 
$\sigma_i^{\textrm{pd}}$ 
is obtained from the central theoretical prediction 
$\sigma_i^{\textrm{th}}$ 
for each bin $i$ out of $N_{\textrm{bins}}$ by varying it with the statistical uncertainties 
$\delta_i^{\textrm{stat}}$, 
obtained in the procedure desribed in Section~\ref{sec:pseudo-data_generation}, 
as 
%TODO once more complete uncertainty estimates are there, add them to the variation
$\sigma_i^{\textrm{pd}} = \sigma_i^{\textrm{th}}(1 + r_i \delta_i^{\textrm{stat}})$, 
with $r_i$ a univariate Gaussian random number.
%
The correlated uncertainties for the pseudodata and the theoretical prediction 
are contained in the nuisance parameter vectors $b^{\textrm{pd}}$ and $b^{\textrm{th}}$, respectively, and the uncorrelated uncertainties in $\Delta_i$.
%
Their effect on $\sigma^{\textrm{th}}$ and $\sigma^{\textrm{pd}}$
is described by the matrices $\Gamma_i^{\textrm{pd}}$ and $\Gamma_i^{\textrm{th}}$.
The indices $\alpha$ and $\beta$ then run over the uncertainty nuisance parameters for the pseudodata and the theoretical prediction, respectively.
%
The nuisance parameter values $b^{\textrm{th(min)}}$ that minimize Eq.~\eqref{eq:profilingchi2} give the central PDFs $f'_0$ optimized to the profiled dataset in the form
\begin{equation}
f_0' = f_0
      + \sum_\beta b_\beta^{\textrm{th(min)}} 
        \left(  \frac{f_\beta^+   -  f_\beta^- }{2}
              -    b_\beta^{\textrm{th(min)}}
                \frac{f_\beta^+ + f_\beta^- - 2f_0}{2}
        \right),
\end{equation}
where $f_0$ is the original central PDF and the up and down variation eigenvectors are given by $f^+, f^-$. The decrease in the uncertainties of the shifted PDFs indicates the enhancement that including the FPF data in the global PDF fit could bring. Here, the profiling studies are performed using version 2.2.1 of the \textsc{xFitter} open-source QCD analysis framework~\cite{Alekhin:2014irh, Bertone:2017tig, xFitter:2022zjb, xFitter:web}, extended to support the use of \textsc{PineAPPL} grids~\cite{christopher_schwan_2023_7499507, Carrazza:2020gss}.
%
Using the PDF4LHC21~\cite{PDF4LHCWorkingGroup:2022cjn} PDF set, the results are illustrated by comparing the fractional uncertainties of the original PDF set to those obtained after profiling. 
This is shown for the valence quark, gluon and the singlet distributions in Fig.~\ref{FASERv2_profiling} for FASER$\nu$2 pseudodata, while Fig.~\ref{FASERv_profiling} indicates the impact of FASER$\nu$. FASER$\nu$2 is observed to provide tighter constraints on the PDFs, with the effects being the most prominent for the valence and sea quark distributions.

>>>>

\begin{center}
  {\Large \bf The LHC as a Neutrino-Ion Collider }\\
  \vspace{1.1cm}
  {\small
Max Fieg$^{1}$, Tommaso Giani$^{2,3}$, Peter Krack$^{2,3}$  Giacomo Magni$^{2,3}$, \\[0.1cm] Toni Makela$^{4}$,
    Tanjona Rabemananjara$^{2,3}$, and Juan Rojo$^{2,3}$, 
  }\\

{\it \small
	~$^1$ UC Irvine  \\[0.1cm]
 	~$^2$Department of Physics and Astronomy, Vrije Universiteit, NL-1081 HV Amsterdam\\[0.1cm]
  	~$^3$Nikhef Theory Group, Science Park 105, 1098 XG Amsterdam, The Netherlands\\[0.1cm]
        ~$^4$ IPJ \\[0.1cm]
 }

@@

\end{center}

Proton-proton collisions at the LHC produce a collimated beam large flux of high-energy neutrinos,
which until recently escaped detection and hence we not used for physics.
%
Detection and analysis of these far-forward LHC neutrinos is taking now place at the
FASER$\nu$ and SND@LHC experiments, and would be uniquely scrutinized
at the proposed Forward Physics Facility (FPF).
%
In this work we demonstrate that exploiting this high-energy far-forward neutrino beam
provides new constraints on the quark and gluon substructure of nucleons and nuclei.
%
We generate projections for inclusive and charm (dimuon)
deep-inelastic scattering (DIS) structure functions
at cross-sections for all ongoing and planned far-forward LHC neutrinos experiments,
including the expected systematic uncertainties.
%
By means of both the xFitter and NNPDF frameworks we quantify the impact of the FPF
measurements on proton and nuclear structure, showing stringent constraints specially
in quark flavour separation.
%
Our analysis demonstrates that instrumenting the far-forward region to exploit the HL-LHC
neutrino fluxes effectively realises a  Neutrino-Ion Collider with
center of mass  energy around 100 GeV at CERN, fully complementary
and operating concurrently to the
Electron-Ion Collider in the US.

@@

We start with a perfect detector.

For bins in $x,Q^2,E_{\nu}$, we start with the \href{https://github.com/juanrojochacon/FPF-WG1/blob/main/results/diff_xsecs_a1.txt}{differential cross section},$\frac{d^2\sigma(x,Q^2,E_{\nu}}{dxdQ^2})$ [pb GeV$^{-2}$] as output from {\tt YADISM} and convolve with the \href{https://github.com/KlingFelix/FastNeutrinoFluxSimulation/tree/main/Fluxes}{neutrino flux}, $\frac{dN_{\nu}}{dE_{\nu}}$. So for a detector with $n_T$ nuclear target density and $L_T$ length, we calculate the event rate per bin as

@@

The statistical uncertainty for each bin is then defined by a Gaussian distribution with fractional uncertainty $1/\sqrt{N{\rm ev}/{\rm bin}}$, i.e. $\delta N_{\rm ev}/{\rm bin} = \sqrt{N_{\rm ev}} /{\rm bin}$

  Following this prescription, we can produce plots which characterize the event rate for a particular neutrino species at a particular experiment.

Results for FASER$\nu$2 can be found \href{https://github.com/juanrojochacon/FPF-WG1/tree/main/results}{here}. The .txt files will display which neutrino species was considered, and also compile experimental geometries and target details in the header. In the sub-folder \href{https://github.com/juanrojochacon/FPF-WG1/tree/main/results/Summaries}{summaries}, plots can be found which show the event rate in $x,Q^2$ space, integrating over the entire neutrino spectrum. The event rate for muon neutrinos at FASER$\nu$2 is shown in Fig~\ref{fig:fasernu2_muon}.

@@

We first generate a MC dataset, $D_0$, over this space with $N = 10^7$ samples and calculate $x,Q^2,E_{\nu}$ for each event, removing samples with $y > 1$. We then integrate according to Eq. \ref{MCintegration} to produce a distribution of events with bins in $x,Q^2,E_{\nu}$. We then smear each sample according to a Gaussian distribution with widths given by Eq.~\ref{fasernu2systematic_errors} to produce a new data set $D_i$ and repeat to produce $M = 10$ event distributions($M>10$ can later be used, but $M = 10$ appears to be stable). For each bin we take the standard deviation to produce the systematic errors per bin (denoted as "N\textunderscore sys\textunderscore errs" in files named "binned\textunderscore sys-events... .txt" \href{https://github.com/juanrojochacon/FPF-WG1/tree/main/results}{here}).

For FASER${\nu}$2, the systematic errors typically dominate over statistical errors, at about $10\%$.

\begin{figure}[h]
    \centering
    \includegraphics[width = 1\textwidth]{error_plot_FASERv2_14.pdf}
    \caption{Event rates with error bars at FASER$\nu$2 for $\nu_{\mu} N \rightarrow \mu^- X$} summed over $Q^2$. The error bars along the x-axis showing the width of the $x$-bins, and are not showing any type of error.
    \label{fig:my_label}
\end{figure}

@@

The expected impact of the FPF data on global PDF fits is assessed by a profiling procedure~\cite{Paukkunen:2014zia, Schmidt:2018hvu, AbdulKhalek:2018rok, HERAFitterdevelopersTeam:2015cre}, based on minimizing the function
%TODO correlations are also included here, as we hopefully get there in the end -- if not, remove terms involving correlations+pseudodata
\begin{equation}
\chi^2 = 
\sum_{i=1}^{N_{\textrm{bins}}} 
\frac{\left(  \sigma_i^{\textrm{pd}}
            + \Gamma_i^{\alpha,\textrm{pd}}
              b_\alpha^{\textrm{pd}}
            - \sigma_i^{\textrm{th}}
            - \Gamma_i^{\beta,\textrm{th}}
              b_\beta^{\textrm{th}}
     \right)^2
     }{\Delta_i^2}
+ \sum_\alpha (b_\alpha^{\textrm{pd}})^2
+ \sum_\beta  (b_\beta^{\textrm{th}})^2 \, .
\label{eq:profilingchi2}
\end{equation}
Here the pseudodata 
$\sigma_i^{\textrm{pd}}$ 
is obtained from the central theoretical prediction 
$\sigma_i^{\textrm{th}}$ 
for each bin $i$ out of $N_{\textrm{bins}}$ by varying it with the statistical uncertainties 
$\delta_i^{\textrm{stat}}$, 
obtained in the procedure described in Section~\ref{sec:pseudo-data_generation}, 
as 
%TODO once more complete uncertainty estimates are there, add them to the variation
$\sigma_i^{\textrm{pd}} = \sigma_i^{\textrm{th}}(1 + r_i \delta_i^{\textrm{stat}})$, 
with $r_i$ a univariate Gaussian random number.
%
The correlated uncertainties for the pseudodata and the theoretical prediction 
are contained in the nuisance parameter vectors $b^{\textrm{pd}}$ and $b^{\textrm{th}}$, respectively, and the uncorrelated uncertainties in $\Delta_i$.
%
Their effect on $\sigma^{\textrm{th}}$ and $\sigma^{\textrm{pd}}$
is described by the matrices $\Gamma_i^{\textrm{pd}}$ and $\Gamma_i^{\textrm{th}}$.
The indices $\alpha$ and $\beta$ then run over the uncertainty nuisance parameters for the pseudodata and the theoretical prediction, respectively.
%
The nuisance parameter values $b^{\textrm{th(min)}}$ that minimize Eq.~\eqref{eq:profilingchi2} give the central PDFs $f'_0$ optimized to the profiled dataset in the form
\begin{equation}
f_0' = f_0
      + \sum_\beta b_\beta^{\textrm{th(min)}} 
        \left(  \frac{f_\beta^+   -  f_\beta^- }{2}
              -    b_\beta^{\textrm{th(min)}}
                \frac{f_\beta^+ + f_\beta^- - 2f_0}{2}
        \right),
\end{equation}
where $f_0$ is the original central PDF and the up and down variation eigenvectors are given by $f^+, f^-$. The decrease in the uncertainties of the shifted PDFs indicates the enhancement that including the FPF data in the global PDF fit could bring. Here, the profiling studies are performed using version 2.2.1 of the \textsc{xFitter} open-source QCD analysis framework~\cite{Alekhin:2014irh, Bertone:2017tig, xFitter:2022zjb, xFitter:web}, extended to support the use of \textsc{PineAPPL} grids~\cite{christopher_schwan_2023_7499507, Carrazza:2020gss}.
%
Using the PDF4LHC21~\cite{PDF4LHCWorkingGroup:2022cjn} PDF set, the results are illustrated by comparing the fractional uncertainties of the original PDF set to those obtained after profiling. 
This is shown for the valence quark, gluon and the singlet distributions in Fig.~\ref{FASERv2_profiling} for FASER$\nu$2 pseudodata, while Fig.~\ref{FASERv_profiling} indicates the impact of FASER$\nu$. FASER$\nu$2 is observed to provide tighter constraints on the PDFs, with the effects being the most prominent for the valence and sea quark distributions.
